---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# checkstring

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/m-muecke/checkstring/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/m-muecke/checkstring/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/checkstring)](https://CRAN.R-project.org/package=checkstring)
<!-- badges: end -->

The goal of checkstring is to provide string format validation functions for R,
inspired by [Zod's](https://zod.dev/?id=strings) string validators.
It extends the [checkmate](https://mllg.github.io/checkmate/) family of
argument checks with common string format validations such as email addresses,
UUIDs, URLs, IP addresses, and financial and academic identifiers like ISIN,
CUSIP, IBAN, ISBN, and ORCID, including check digit verification.

## Installation

You can install the released version of **checkstring** from [CRAN](https://CRAN.R-project.org) with:

```{r}
#| eval: false
install.packages("checkstring")
```

And the development version from [GitHub](https://github.com/) with:

```{r}
#| eval: false
# install.packages("pak")
pak::pak("m-muecke/checkstring")
```

## Usage

```{r}
library(checkstring)

is_email("user@example.com")
is_uuid("550e8400-e29b-41d4-a716-446655440000")
is_ipv4("192.168.1.1")
is_hex("deadbeef")
is_url("https://example.com")
is_base64("SGVsbG8gV29ybGQ=")
is_isin("US0378331005")
is_iban("GB29NWBK60161331926819")
```

## Related work

* [checkmate](https://github.com/mllg/checkmate): Fast and versatile argument checks for R
* [figir](https://github.com/philaris/figir): Check validity of FIGI, CUSIP, ISIN, SEDOL
* [libbib](https://github.com/NYPL/libbib): Validate and normalize ISBN, ISSN, LCCN
